---
title: "Data analysis code for IOG Experiment"
output: html_document
date: "2024-02-15"
---

# Load packages and directories

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(tidyr)
library(multcomp)
library(knitr)
library(ggplot2)
library(broom)
library(rcompanion)
library(ggpubr)
library(plotrix)
library(gridExtra)
library(dotwhisker)
library(ez)
library(car)
```

# Specify the path to the rawdata folder

```{r}
project_dir <- file.path('~','Documents','Interocular-grouping-and-motion')
rawdata_path <- file.path(project_dir, "rawdata")
```

# Specify the subjects

```{r}
subjects <- c('sub-01', 'sub-02', 'sub-03', 'sub-04', 'sub-05', 'sub-06', 'sub-07', 'sub-08', 'sub-09')
```

# Create an empty data frame

```{r}
mergedData <- data.frame()
```

# Loop through all csv files for each subject and merge them into one data frame

```{r}
for (subject in subjects) {
  # Create the path to the subject folder
  subjectFolderPath <- file.path(rawdata_path, subject)
  
  # List CSV files in the subject folder
  files <- list.files(subjectFolderPath, pattern = "*.csv", full.names = TRUE)
  # get list of files for experiment 1
  files <- files[grepl("task-IOG_run", files)]
  # excluded files containing 'conditions' in name
  files <- files[!grepl("conditions", files)]
  
  # Read and merge CSV files
  run <- 1
  for (file in files) {
    data <- read.csv(file)
    data$subject = subject
    data$run = run
    run <- run+1
    mergedData <- rbind(mergedData, data)
  }
}
```


# Anova and post-hoc tests for proportion of mixed percepts across different conditions

```{r}
mergedData <- aggregate(durations ~ condition + percepts + subject,
                         data = mergedData,
                       FUN = sum)

mergedData$condition <- factor(mergedData$condition)
mergedData$subject <- factor(mergedData$subject)
```

```{r}
mergedData_mixed <- mergedData %>%
  group_by(subject, condition) %>%
  summarise(proportion_mixed = sum(durations[percepts == 'mixed'])/sum(durations),
    .groups = 'drop')

aov_mixed <- ezANOVA(data = mergedData_mixed,
                   dv = proportion_mixed,
                   within = condition,
                   wid = subject)
print(aov_mixed)
```

## Post-hoc t-tests
```{r}
NMNC <- mergedData_mixed$proportion_mixed[mergedData_mixed$condition == 'NoMotionNoColor']
MNC <- mergedData_mixed$proportion_mixed[mergedData_mixed$condition == 'MotionNoColor']
NMC <- mergedData_mixed$proportion_mixed[mergedData_mixed$condition == 'NoMotionColor']
MC <- mergedData_mixed$proportion_mixed[mergedData_mixed$condition == 'MotionColor']
```

```{r}
# main tests
motion_cue_res <- t.test(x = NMNC, y = MNC, alternative = 'two.sided')  # expectation significant difference
print(motion_cue_res)
color_cue_res <- t.test(x = NMNC, y = NMC, alternative = 'two.sided')   # expectation significant difference
print(color_cue_res)
motion_vs_color <- t.test(x = MNC, y = NMC, alternative = 'two.sided') # expectation NO significant difference
print(motion_vs_color)
motion_color_vs_motion <- t.test(x = MNC, y = MC, alternative = 'two.sided')  # expectation significant difference
print(motion_color_vs_motion)
motion_color_vs_color <- t.test(x = NMC, y = MC, alternative = 'two.sided')  # expectation significant difference
print(motion_color_vs_color)
motion_color_vs_none <- t.test(x = NMNC, y = MC, alternative = 'two.sided')  # expectation significant difference
print(motion_color_vs_none)
```


```{r}
mergedData$condition <- factor(mergedData$condition, levels = c('NoMotionNoColor', 'MotionNoColor', 'NoMotionColor', 'MotionColor'))
```

# Bar Plot of Mixed Proportions across conditions
```{r}
bar_plot_mixed <- ggplot(data = mergedData_mixed, aes(x = condition, 
                                                      y = proportion_mixed, 
                                                      fill = condition
                                                      )
                         ) +
  geom_bar(stat = "summary", 
           position = "dodge") +
  labs(x = "Conditions", y = "Proportion of 'Mixed' Percepts") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 0.4))

print(bar_plot_mixed)
```
# Filtering data for proportion interocular without mixed percepts

```{r}
# Filter the data to exclude rows with 'mixed' percepts
filteredData <- mergedData[mergedData$percepts != 'mixed', ]

# Convert 'percepts' column to a factor with levels 'interocular', 'monocular'
filteredData$percepts <- factor(filteredData$percepts, levels = c('interocular', 'monocular'))

durations <- filteredData$durations

filteredData$condition <- factor(filteredData$condition, levels = c('NoMotionNoColor', 'MotionNoColor', 'NoMotionColor', 'MotionColor'))

percepts <- unique(filteredData$percepts)

# Aggregate dominance durations by subject, typical color and percept using sum
# instead of median

filteredData <- aggregate(durations ~ condition + percepts + subject,
                         data = filteredData,

                       FUN = sum)
filteredData_IOG <- filteredData %>%
group_by(subject, condition) %>%
summarise(proportion_IOG = (
sum(durations[percepts == 'interocular']))/(sum(durations[percepts == 'interocular']+
                                                durations[percepts == 'monocular'])),
.groups = 'drop')

```

# Bar plot for Proportion IOG without mixed
```{r}
bar_plot_IOG <- ggplot(data = filteredData_IOG, aes(x = condition, y = proportion_IOG, fill = condition)) +
  geom_bar(stat = "summary", position = "dodge") +
 labs(x = "Conditions", y = "Proportion of 'Interocular' Percepts") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 1))

print(bar_plot_IOG)
```

# Linear regression test
```{r}
filteredData_IOG$condition_numeric <- ifelse(filteredData_IOG$condition == "NoMotionNoColor", 0,
                                       ifelse(filteredData_IOG$condition %in% c('NoMotionColor', 'MotionNoColor'), 1, 2))

# Perform linear regression
lm_test_IOG <- lm(proportion_IOG ~ condition_numeric, data = filteredData_IOG)

residuals <- residuals(lm_test_IOG)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

print(shapiro_test) # Shapiro test's p-value is .3324, meaning residuals of the data do not
                    # significantly deviate from normal and the model is a good fit for the data


summary(lm_test_IOG)
```
```{r}
hist(residuals)
```


```{r}
qq1 <- qqPlot(residuals)

```

```{r}
filteredData_IOG$subject <- factor(filteredData_IOG$subject)

# ANOVA test
aov_IOG <- ezANOVA(data = filteredData_IOG,
                   dv = proportion_IOG,
                   within = condition,
                   wid = subject)
print(aov_IOG)
```

## Post-hoc t-tests
```{r}
NMNC <- filteredData_IOG$proportion_IOG[filteredData_IOG$condition == 'NoMotionNoColor']
MNC <- filteredData_IOG$proportion_IOG[filteredData_IOG$condition == 'MotionNoColor']
NMC <- filteredData_IOG$proportion_IOG[filteredData_IOG$condition == 'NoMotionColor']
MC <- filteredData_IOG$proportion_IOG[filteredData_IOG$condition == 'MotionColor']
```

```{r}
# main tests
motion_cue_res <- t.test(x = NMNC, y = MNC, alternative = 'less')  # expectation significant difference
print(motion_cue_res)
color_cue_res <- t.test(x = NMNC, y = NMC, alternative = 'less')   # expectation significant difference
print(color_cue_res)
motion_vs_color <- t.test(x = MNC, y = NMC, alternative = 'two.sided') # expectation NO significant difference
print(motion_vs_color)
motion_color_vs_motion <- t.test(x = MNC, y = MC, alternative = 'less')  # expectation significant difference
print(motion_color_vs_motion)
motion_color_vs_color <- t.test(x = NMC, y = MC, alternative = 'less')  # expectation significant difference
print(motion_color_vs_color)
```

# Bar plot of proportion IOG

```{r}
# Create bar plot of mean proportions of 'Interocular' percepts
bar_plot_interocular <- ggplot(data = filteredData_IOG, aes(x = condition, y = proportion_IOG, fill = condition)) +
  geom_bar(stat = "summary", position = "dodge") +
  labs(x = "Conditions", y = "Proportion of 'Interocular' Percepts") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 1.1)) +
  geom_segment(aes(x = 1.0, xend = 3.0, y = 1.0, yend = 1.0), size = 0.5) +
  geom_segment(aes(x = 1.0, xend = 1.0, y = 0.99, yend = 1.0), size = 0.5) +
  geom_segment(aes(x = 3.0, xend = 3.0, y = 0.99, yend = 1.0), size = 0.5) +
  geom_segment(aes(x = 2.3, xend = 2.9, y = 0.92, yend = 0.92), size = 0.5) +
  geom_segment(aes(x = 2.3, xend = 2.3, y = 0.91, yend = 0.92), size = 0.5) +
  geom_segment(aes(x = 2.9, xend = 2.9, y = 0.91, yend = 0.92), size = 0.5) +
  geom_segment(aes(x = 2.0, xend = 4.0, y = 0.82, yend = 0.82), size = 0.5) +
  geom_segment(aes(x = 2.0, xend = 2.0, y = 0.81, yend = 0.82), size = 0.5) +
  geom_segment(aes(x = 4.0, xend = 4.0, y = 0.81, yend = 0.82), size = 0.5) +
  geom_text(aes(x = 2.0, y = 0.98, label = "*"), size = 4, vjust = -1) +
  geom_text(aes(x = 2.50, y = 0.9, label = "*"), size = 4, vjust = -1) +
  geom_text(aes(x = 2.60, y = 0.9, label = "*"), size = 4, vjust = -1) +
  geom_text(aes(x = 2.70, y = 0.9, label = "*"), size = 4, vjust = -1) +
  geom_text(aes(x = 3.0, y = 0.79, label = "*"), size = 4, vjust = -1) +
  geom_text(aes(x = 3.1, y = 0.79, label = "*"), size = 4, vjust = -1)

# Print the plot
print(bar_plot_interocular)
```
